From Freebase to Wikidata: The Great Migration

### 2016-01-29 paper.tex
Reworked Section 5.3.
Reworked Section 6.

@@

Our theory is that this is
\emph{(i)}~due to the absence of a~lot of musical data in Wikidata that is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.

The big horizontal lines in \autoref{fig:freebase} and even more in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.

@@

From the Freebase dump, we have been able to create more than 17~million Wikidata claims
(\emph{i.e.}, statements without references), including 1.5~million IDs.
For that, we have used 31~million Freebase facts.
Including the references, we obtain 19.6~million statements, and, after removing duplicates
and facts already contained in Wikidata, we obtain 14~million new statements.

If all these statements were added to Wikidata,
we would see a~21\% increase of the number of statements in Wikidata.

This raises the question why these numbers are so low compared to the size of Freebase, given at
3~billion facts.
The main reason is that we have only mapped 4.56~million items to Wikidata,
\emph{i.e.}, only 9.5\% of Freebase's topics in total.
The mapped topics are the subject of only 64~million facts.
So even under otherwise ideal circumstances, we cannot map more than 64~million statements.
This assumes that we could map all reverse properties, which, in fact, we could not.
So leaving aside the not mapped topics, we have created a~statement for more than 24\% of the facts.

@@

For the challenges listed in \autoref{sec:challenges-of-the-migration}
there is a~body of work dealing with it. In many cases, the problems we have encountered in
this migration have turned out to be much simpler than the problems solved in existing research.

\textbf{Data licensing.}
Although many books and tutorials on publishing data on the Web of Data
(e.g.~\cite{linkeddatabook,linkeddatabestpractises}) mention the importance of licenses,
they are often added as an afterthought.
They have been preciously few dedicated publications on the
topic~\cite{licenses,licensius}, none of them written by lawyers.
Due to the lack of understanding surronding these topics, we have been
cautious and chose to simply remove any facts from the migrating data
where the situation is in doubt.

\textbf{Schema mappings.}
There has been extensive research on the topic of mapping the structures
of two databases or knowledge bases
(some surveys are~\cite{mappingsurvey1,mappingsurvey2,mappingsurvey3}).
The current work is, in many technical respects, rather simple:
we are looking for a~one-directional mapping (since the case of data
flowing from Wikidata back to Freebase is not relevant),
a~one-off data flow (since Freebase is now read-only),
mostly individually evaluated by human contributors
(therefore not requiring a~perfect mapping).
The effort most similar to ours is the DBpedia Mapping Wiki~\cite{lehmann2014dbpedia}.
The DBpedia community is manually writing and maintaining the mapping from
the DBpedia extraction process (working on Wikipedia) to the DBpedia ontology.
The main difference is that the DBpedia mapping is running continuously
and does not allow for point-fixes, thus necessitating a~higher quality of the mapping,
whereas for the Freebase to Wikidata mapping every single results is expected to be curated.
The schema.org structured data project is also exploring mappings to and from Wikidata,
to facilitate the use of Wikidata's vocabulary as schema.org markup across the wider Web.
This involves similar issues mapping schema.org's CVT-like statement qualification mechanism, Roles,
to Wikidata conventions.%
\footnote{On Roles: \url{http://schema.org/Role}. Mappings from Schema.org to Wikidata:
\url{https://github.com/schemaorg/schemaorg/issues/280}. On the interaction of Schema.org
and Wikidata: \url{https://meta.wikimedia.org/wiki/Wikidata/Notes/Schema.org_and_Wikidata}}

\textbf{Entity mappings.}
Entity mapping (often called entity linking) deals with finding the objects
in several sources that refer to the same entity in the domain of discourse.
There is extensive work in this area (e.g.~\cite{entitylinkingsurvey1,entitylinkingsurvey2})
but we avoided most of the challenges by relying on the simple fact that
both Freebase and Wikidata have historically been bootstrapped by using Wikipedia.
Wikipedia is often used as a~benchmark for entity mapping systems~\cite{entitylinkingwikipedia}.
As described in \autoref{subsec:topic-mappings} this lead to sufficiently good results for the head,
and we think it would be surprising if further effort in this area would lead to reasonable gains.

\textbf{Data maintenance.}
The Wikidata requirements~\cite{wikidatareqs} states that the data uploaded to Wikidata
must not overwhelm the community.
Research in understanding how systems for collaborative knowledge creation are impacted
by events like this data migration are still in its early stages~\cite{moskaliuk2012collaborative},
in particular for structured knowledge~\cite{horridge2014webprotege}.
Most of the research is focused on Wikipedia~\cite{flock2015towards},
which is understandable considering the availability of its data sets,
in particular the whole edit history~\cite{schindler2011introducing}
and the availability of tools for working with Wikipedia~\cite{milne2013open}.

>>>>

\hyphenation{DBpedia Wikipedia Wikidata Freebase}

@@

Our theory is that this is
\emph{(i)}~due to the absence of a~lot of musical data in Wikidata that is present in Freebase,
and \emph{(ii)} due to the US population CVTs in Freebase.
The big horizontal lines in \autoref{fig:freebase} and even more in \autoref{fig:wikidata}
can most probably be traced back to places that do not have any linked date,
and in consequence inherit the foundation date of their country.
For Wikidata (\autoref{fig:wikidata}), two lines that are definitely caused by this effect
are the big line over the USA around 1780
and the one covering Russia around 850.

@@

From the Freebase dump, we have been able to create more than 17~million Wikidata claims
(\emph{i.e.}, statements without references), including 1.5~million IDs.
For that, we have used 31~million Freebase facts.
Including the references, we obtain 19.6~million statements, and, after removing duplicates
and facts already contained in Wikidata, we obtain 14~million new statements.
If all these statements were added to Wikidata,
we would see a~21\% increase of the number of statements in Wikidata.

This raises the question why these numbers are so low compared to the size of Freebase of 3~billion facts.
The main reason is that we have only mapped 4.56~million items to Wikidata,
\emph{i.e.}, only 9.5\% of Freebase's topics in total.
The mapped topics are the subject of only 64~million facts.
So even under otherwise ideal circumstances, we cannot map more than 64~million statements.
This assumes that we could map all reverse properties, which, in fact, we could not.
So leaving aside the not mapped topics, we have created a~statement for more than 24\% of the facts.

@@

For the challenges listed in \autoref{sec:challenges-of-the-migration},
there are a~number of works dealing with it. In many cases, the problems we have encountered in
this migration have turned out to be much simpler than the problems solved in existing research.

\textbf{Data licensing.}
Although many books and tutorials on publishing data on the Web of Data
(e.g.~\cite{linkeddatabook,linkeddatabestpractises}) mention the importance of licenses,
they are often added as an afterthought.
There have been remarkably few dedicated publications on the
topic~\cite{licenses,licensius}, none of them written by lawyers.
Due to the lack of understanding surrounding these topics, we have been
cautious and chose to simply remove any facts from the to-be-migrated data
if there were doubts.

\textbf{Schema mappings.}
There has been extensive research on the topic of mapping the structures
of two databases or knowledge bases
(some surveys are~\cite{mappingsurvey3,mappingsurvey2,mappingsurvey1}).
The current work is, in many technical respects, rather simple:
we are looking for a~one-directional mapping (since the case of data
flowing from Wikidata back to Freebase is not relevant),
a~one-off data flow (since Freebase is now read-only),
mostly individually evaluated by human contributors
(therefore not requiring a~perfect mapping).
The effort most similar to ours is the DBpedia Mapping Wiki~\cite{lehmann2014dbpedia}.
The DBpedia community is manually writing and maintaining the mapping from
the DBpedia extraction process (working on Wikipedia) to the DBpedia ontology.
The main difference is that the DBpedia mapping is running continuously
and does not allow for point-fixes, thus necessitating a~higher quality of the mapping,
whereas for the Freebase to Wikidata mapping every single result is expected to be curated.
The Schema.org structured data project is also exploring mappings to and from Wikidata,
to facilitate the use of Wikidata's vocabulary as Schema.org markup across the wider Web.
This involves similar issues mapping\linebreak Schema.org's CVT-like statement qualification mechanism and roles
to Wikidata conventions.%
\footnote{On Roles: \url{http://schema.org/Role}. Mappings from Schema.org to Wikidata:
\url{https://github.com/schemaorg/schemaorg/issues/280}. On the interaction of Schema.org
and Wikidata: \url{https://meta.wikimedia.org/wiki/Wikidata/Notes/Schema.org_and_Wikidata}}

\textbf{Entity mappings.}
Entity mapping (often called entity linking) deals with finding the objects
in several sources that refer to the same entity in the domain of discourse.
There is extensive work in this area (e.g.~\cite{entitylinkingsurvey1,entitylinkingsurvey2})
but we avoided most of the challenges by relying on the simple fact that
both Freebase and Wikidata have historically been bootstrapped by using Wikipedia.
Wikipedia is often used as a~benchmark for entity mapping systems~\cite{entitylinkingwikipedia}.
As described in \autoref{subsec:topic-mappings} this leads to sufficiently good results,
and we argue it would be surprising if further effort in this area would lead to reasonable gains.

\textbf{Data maintenance.}
The Wikidata requirements~\cite{wikidatareqs} state that the data uploaded to Wikidata
must not overwhelm the community.
Research in understanding how systems for collaborative knowledge creation are impacted
by events like this data migration are still in its early stages~\cite{moskaliuk2012collaborative},
in particular for structured knowledge~\cite{horridge2014webprotege}.
Most of the research is focused on Wikipedia~\cite{flock2015towards},
which is understandable considering the availability of its data sets,
in particular the whole edit history~\cite{schindler2011introducing}
and the availability of tools for working with Wikipedia~\cite{milne2013open}.
